### 2017-01-26 papers/outline.tex
Adds paper outline

\documentclass{article}

\title{Progtran Paper Outline}
\date{01-26-2017}
\author{Lucas Saldyt}

\begin{document}

\maketitle
\pagenumbering{gobble}
\newpage
\pagenumbering{arabic}

\section{Abstract}
Progtran is an attempt to build a multi-language source-to-source compiler. Ideally, one will be able to convert programs in one language to programs in another without a loss in readability. This will allows for more user flexibility. For example, Python syntax is generally easier to understand than C++'s, but if one can translate Python to C++, then one can have both benefits. Progtran is an investigation into the feasibility of such conversions.
\section{Introduction}
\section{Methods}
\subsection{Grammars}
\subsection{Constructors}
\subsection{AST Transformations}
\section{Results}

\end{document}